taxize - taxonomic search and retrieval in R

### 2013-07-05 inst/doc/taxize_withcode_plos.Rnw
replaced aggregation example

\subsection*{Aggregating data by any taxonomic rank}

In biology, one can asks questions at varying taxonomic levels. Do species differ? Do genera differ? This use case is easily handled in taxize. A very basic function called \emph{rankagg} will aggregate any data to a specific taxonomic level. In this example, a single column \emph{abundance} is aggregated by any arithmetic function (e.g., mean) at any taxonomic rank (e.g., Family).

<<aggregate_abd, eval=TRUE, cache=TRUE>>=
library(vegan)
data(dune.taxon)
dat <- dune.taxon
dat$abundance <- round(rlnorm(n = nrow(dat), meanlog = 5, sdlog = 2), 0)
head(dat)
head( rankagg(data = dat, datacol = "abundance", rank = "Genus") )
@

>>>>

\subsection*{Aggregating data to a specific taxonomic rank}

In biology, one can asks questions at varying taxonomic levels. One may perform analyses on different taxonomic levels. This use case is easily handled in taxize. A function called \emph{tax\_agg} will aggregate community data to a specific taxonomic level. In this example, we take data of 5 species and aggregate them to family level. Again we can specify if we want to use data from ITIS or NCBI.

<<tax_agg ,eval=TRUE, cache=TRUE, tidy=FALSE>>=
data(dune, package = 'vegan')
df <- dune[ , 1:5]
colnames(df) <- c("Bellis perennis", "Empetrum nigrum", "Juncus bufonius", 
                  "Juncus articulatus", "xxx")
head(df)
agg <- tax_agg(df, rank = 'family', db = 'ncbi')
agg
head(agg$x)
@

We see that the two Juncus species are aggregated to Juncaceae and their abundances are summed up. If a taxon is one lower taxonomic resolution then the querried or the taxon is not found in the database (like in the example taxon \emph{xxx}) then these are not aggregated and returned as is.
